---
title: "analyze_results_bp"
author: "Ben"
date: "7/2/2018"
output: html_document
---

# Notes

  * Relation between initital `t_prob` and `B_prob` and eventual stability
  * Few (0.13) of simulations result in optional stability...

# Interperation

### Recursion vs generational change interpretation?
  * Anything significant about this interpretation?



```{r libraries, message=FALSE, warning=FALSE}
library(knitr)
library(tidyverse)
```

```{r read-data}
f_p <- 'results-100.csv'
df_results <- read.csv(file.path(getwd(), 'outputs', f_p))
df_results <- df_results %>%
  mutate(is_final_generation=ifelse(is_final_generation=="True", TRUE, FALSE))
```

```{r report}
cat(paste0(length(unique(df_results$seed))), "simulations run")
```

```{r counting-nas, eval=FALSE, echo=FALSE, message=FALSE}
# The original implementation was leading to lots of NaNs.
# This code was examining that behavior.
knitr::kable(
df_results %>%
  filter(is_final_generation) %>%
  summarise(total=n(),
            cnt_na=sum(is.na(r)),
            prop_na=cnt_na/total))

# NOTE (BP): NAs look related to the `that-rate`. If that-rate goes to 0 or 1 there will be no
# intergenerational change???
# df_results %>%
#   filter(that_rate > 0.001, that_rate < 0.999) %>%
#   summarise(sum(is.na(r)))
```

# Trajectories
```{r stable-results-list}
# Get stable trajectories
stables <- df_results %>% 
  mutate(klass=paste0('k-', k, 'c-', c, 'seed-', seed)) %>%
  filter(is_final_generation) %>%
  mutate(stable=ifelse(that_rate > 0.001 & that_rate < 0.999, TRUE, FALSE)) %>%
  filter(stable)
stables <- stables$klass
```

## `that-rate` trajectories
```{r plot-all-that-rate-trajectories}
# Plot all trajectories  
n_samples <- 100
seed_samples <- sample(df_results$seed, n_samples)
df_results %>%
  filter(seed %in% seed_samples) %>%
  mutate(klass=paste0('k-', k, 'c-', c, 'seed-', seed),
         num_generations=as.numeric(num_generations),
         stable=ifelse(klass %in% stables, 'stable-optionality', 'conventionalized')) %>%
  ggplot(aes(x=num_generations, y=that_rate, col=stable, group=klass)) +
    geom_line(alpha=0.3) +
    theme_classic() +
    theme(legend.position="none") +
    ggtitle("that-rate trajectories")
```

```{r plot-facet-that-rate-trajectories}
# Facet by initial B_prob, t_prob
n_samples <- 10
seed_samples <- sample(df_results$seed, n_samples)
df_results %>%
  filter(seed %in% seed_samples) %>%
  mutate(klass=paste0('k-', k, 'c-', c, 'seed-', seed),
         num_generations=as.numeric(num_generations),
         stable=ifelse(klass %in% stables, 'stable-optionality', 'conventionalized')) %>%
  ggplot(aes(x=num_generations, y=that_rate, col=stable, group=klass)) +
    geom_line(alpha=0.3) +
    theme_classic() +
    theme(legend.position="none") +
    facet_grid(round(t_prob, 3)*round(B_prob, 3)~.) +
    ggtitle(paste0(n_samples, " samples"))
```

## `r` trajectories
```{r plot-all-r-trajectories}
df_results %>%
  mutate(klass=paste0('k-', k, 'c-', c, 'seed-', seed),
         num_generations=as.numeric(num_generations),
         stable=ifelse(klass %in% stables, 'stable-optionality', 'conventionalized')) %>%
  ggplot(aes(x=num_generations, y=r, col=stable, group=klass)) +
    geom_line(alpha=0.3) +
    theme_classic() +
    theme(legend.position="none") +
    ggtitle("r trajectories")
```

```{r plot-facet-r-trajectories}
# Facet by initial B_prob, t_prob
df_results %>%
  filter(seed %in% seed_samples) %>%
  mutate(klass=paste0('k-', k, 'c-', c, 'seed-', seed),
         num_generations=as.numeric(num_generations),
         stable=ifelse(klass %in% stables, 'stable-optionality', 'conventionalized')) %>%
  ggplot(aes(x=num_generations, y=r, col=stable, group=klass)) +
    geom_line(alpha=0.3) +
    theme_classic() +
    theme(legend.position="none") +
    facet_grid(round(t_prob, 3)*round(B_prob, 3)~.)
```


# Reproducing R. Levy plots

## Preprocessing
```{r}
# R. Levy's preprocessing
# dat$stable <- with(dat,thatrate > 0.001 & thatrate < 0.999)
# dat <- subset(dat, ! (k==1.0 & c==0.0))
# dat.summary <- dat %>% group_by(k,c) %>%
#   dplyr:::summarise(stable=mean(stable),r=mean(r))

# This preprocessing is only valid for stable optionality plot
df_preprocessed <- df_results %>%
  filter(is_final_generation) %>%
  mutate(stable = that_rate > 0.001 & that_rate < 0.999) %>%
  filter(k != 1.0, c != 0.0) %>%
  group_by(k, c) %>%
  summarise(stable=mean(stable), r=mean(r))
```

## Stable optionality plot
```{r plot-reproduce-stable-optionality-grid}
df_preprocessed %>%
  ggplot(aes(k,c)) + 
    geom_tile(aes(fill=stable),colour="white") +
    labs(y=expression(paste("String length cost parameter ", c)), fill="stable\noptionality\nrate") +
    theme_classic() +
    scale_x_continuous(name=expression(paste("Nonuniformity penalization parameter ",k)),
                       breaks=seq(1,2,by=0.2))
```


## Marginal frequency plot
```{r plot-marginal-frequency}
## Distribution of marginal frequencies of optional marker t at fixed points with stable optionality
df_results %>%
  filter(is_final_generation) %>%
  mutate(stable=that_rate > 0.001 & that_rate < 0.999) %>%
  filter(stable) %>%
  ggplot(aes(x=that_rate,y=..density..)) + 
    geom_histogram(bins=42) + 
    scale_x_continuous(limits=c(-0.05,1.05)) +
    ylab("Probability density") +
    xlab(expression(paste("Marginal frequency of optional marker ",t))) +
    theme_classic()
```

## Perason correlation plot
```{r}
## distribution of correlations between phrase onset probability and t-rate at fixed points with stable optionality
df_results %>%
  filter(is_final_generation) %>%
  mutate(stable = that_rate > 0.001 & that_rate < 0.999) %>%
  filter(stable) %>%
  ggplot(aes(x=r, y=..density..)) + 
    geom_histogram(bins=42) + 
    scale_x_continuous(limits=c(-1.05,1.05)) +
    ylab("Probability density") +
    xlab("Pearson correlation between\nphrase onset & t probabilities") + 
    theme_classic()
```

# Additional analyses
```{r that-rate-r-correlation}
df_results %>% 
  filter(is_final_generation) %>% 
  mutate(stable = that_rate > 0.001 & that_rate < 0.999) %>%
  ggplot(aes(x=that_rate, y=r, col=stable)) +
    geom_point(alpha=0.3) +
    theme_classic()
```

Proportion stable?
```{r proportion-stable}
df_results %>% 
  filter(is_final_generation) %>% 
  mutate(stable = that_rate > 0.001 & that_rate < 0.999) %>%
  group_by(stable) %>%
  summarise(n=n()) %>%
  ungroup %>%
  mutate(total=sum(n),
         prop=n/total) %>%
  ggplot(aes(x=stable, y=prop)) +
    geom_bar(stat='identity') +
    geom_text(aes(x=stable, y=prop, label=round(prop, 2)), nudge_y=0.025) +
    ylim(0, 1) +
    ggtitle("Proportion of simulations leading to stable optionality") +
    theme_classic()
```

Descriptives
```{r num-generations-cnt-stable}
df_results %>% 
  mutate(stable = that_rate > 0.001 & that_rate < 0.999) %>%
  filter(stable) %>%
  select(num_generations) %>%
  summary()
  
```


```{r plot-num-generations-cnt-stable}
df_results %>% 
  mutate(stable = that_rate > 0.001 & that_rate < 0.999) %>%
  filter(stable) %>%
  group_by(num_generations) %>%
  summarise(cnt=n()) %>%
  ggplot(aes(x=num_generations, y=cnt)) +
    geom_bar(stat='identity') +
    theme_classic()
```

```{r log-log-plot-num-generations-cnt-stable}
df_results %>%
  mutate(stable = that_rate > 0.001 & that_rate < 0.999) %>%
  filter(stable) %>%
  mutate(num_generations=num_generations+1) %>%
  group_by(num_generations) %>%
  summarise(cnt=n()) %>%
  ggplot(aes(x=log(num_generations), y=log(cnt))) +
    geom_point(stat='identity', size=2.5, alpha=0.4) +
    geom_smooth(method='lm') +
    theme_classic()
```


Relation between `t_prob` and eventual stability?
```{r plot-relationship-btwn-t-prob-and-stability}
df_results %>%
  filter(is_final_generation) %>%
  mutate(stable = that_rate > 0.001 & that_rate < 0.999) %>%
  filter(stable) %>%
  ggplot(aes(x=num_generations, y=t_prob, col=k, size=c)) +
    geom_point(alpha=0.3) +
    theme(legend.position = 'none') +
    theme_classic()
```
